---
title: "Interactive Visualization"
author: "Lokesh Surendra Jain"
date: '2023-03-17'
output:
  html_document:
    df_print: paged
  pdf_document: default
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, results="hide", warning=F, message=F}
library(data.table)
library(dplyr)
library(dplyr)
library(tidyr)
library(plotly)
library(lubridate)
```


> "The Iris flower data set or Fisher's Iris data set is a multivariate
> data set introduced by the British statistician and biologist Ronald
> Fisher in his 1936 paper The use of multiple measurements in taxonomic
> problems as an example of linear discriminant analysis"
> <https://en.wikipedia.org/wiki/Iris_flower_data_set>

```{r}
iris_data <- fread("C:/Users/lokes/Downloads/iris.csv", stringsAsFactors = TRUE)
# iris_data
```

```{r}
iris_data
```

```{r}

plot_ly(iris_data, x = ~Sepal.Length, color = ~Species, type = "histogram") %>%
  layout(xaxis = list(title = "Sepal Length", tickvals = c(4, 5, 6, 7, 8)), yaxis = list(title = "Count"))
```

```{r}
p <- ggplot(iris_data, aes(x = Sepal.Length, fill = Species)) +
    geom_histogram(binwidth = 0.5, position = "dodge") +
    labs(x = "Sepal Length", y = "Count")

toWebGL(ggplotly(p))
```

```{r} 
iris_long <- iris_data %>% gather(key = "metric", value = "value", -Species)

ggplot(iris_long, aes(x = value, fill = Species)) +
  geom_histogram(bins=30, binwidth = 0.5, position="dodge") +
  facet_wrap(~metric, nrow = 2) +
  labs(x = "value", y = "count")
```

The most effective species separations appear to be between petals'
length and width. Each species' peak is clearly visible in the
histograms for these measurements, with little overlap between them. The
histograms for sepal length and sepal width, on the other hand, exhibit
less distinct demarcation between the species and more overlap.

```{r} 
plot_ly(iris_long, x = ~ value, y = ~ interaction(Species, metric), color = ~ Species) %>%
  add_boxplot()%>%
  layout(title = "Metrics by Species",
         xaxis = list(title = "Value"),
         yaxis = list(title = "interaction(Species, metric)"),
         facet_col = ~Species + metric)

```

```{r} 

plot_ly(iris_data, x = ~Petal.Length, y = ~Petal.Width, color = ~Species, type = "scatter", mode = "markers", showlegend = TRUE) %>%
  layout(title = "Iris Dataset: Petal Length vs. Petal Width",
         xaxis = list(title = "Petal Length"),
         yaxis = list(title = "Petal Width"),
         legend = list(title = "Species") )

```

```{r} 
plot_ly(iris_data, x = ~Petal.Length, y = ~Petal.Width, z = ~Sepal.Length,
        color = ~Species, type = "scatter3d", mode = "markers")
```

We can observe from the scatter plot that there is a distinct division
between the three species. The three species can clearly be seen
clustering together in the Petal Length vs. Petal Width scatter plot,
with Setosa forming a well-separated cluster and Versicolor and
Virginica overlapping. This indicates that while Petal Length and Petal
Width are useful measures for separating Setosa and the other two
species, Versicolor and Virginica may not be as well-served by them. The
Iris dataset is a suitable case study for investigating data
visualization strategies, according to the scatter plot, as it amply
highlights the value of utilizing various metrics to visualize data
trends.
